### 2018-06-11 RA-L/theory.tex
moved epsilon description into assumptions
Changed the image of the algorithm
Added the full algorithm to the paper
Previous changes
Shiva PhD dissertation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
\label{sec:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htb]
\begin{center}
\includegraphics[width=\linewidth]{SquareFill.pdf} 
\vspace{-1em}
\caption{Pushing the swarm against a square boundary wall allows limited control of the shape of the swarm, as a function of swarm area $A$ and the commanded movement direction $\beta$. Left plot shows locus of possible mean positions for five values of $A$.  The locus morphs from a square to a circle as $A$ increases.  The covariance ellipse for each $A$ is shown with a dashed line. Center shows two corresponding arrangements of kilobots.  At right is $\bar{x}(A), \sigma_{xy}(A), \sigma_x^2(A),$ and $\rho(A)$ for a range of $\beta$ values. See online interactive demonstration at \citep{Zhao2016mathematicaSquare}.}
\label{fig:SquareFill}
\end{center}
\end{figure*} 
\begin{figure*}[!htb]
\begin{center}
\vspace{-1em}
\includegraphics[width=\linewidth]{CircleFill.pdf} 
\vspace{-1em}
\caption{Pushing the swarm against a circular boundary wall allows limited control of the shape of the swarm, as a function of the fill level $h$ and the commanded movement direction $\beta$. Left plot shows locus of possible mean positions for four values of $h$. The locus of possible mean positions are concentric circles. See online interactive demonstration at \citep{Zhao2016mathematica}.}  
\label{fig:CircleFill}
\end{center}
\end{figure*}

@@

The next section shows how a system in a rectangularly bounded workspace with friction model \eqref{eq:frictionmodel} can arbitrarily position two robots. 
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\columnwidth]{friction.pdf} 
\vspace{-1em}
\caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (c) velocity of a fluid reduces to zero at the boundary.
\label{fig:friction}
}
\end{center}
\end{figure}

\section{Algorithms}\label{sec:algorithms}
\subsection{Position Control of Two Robots Using Wall Friction}\label{sec:PostionControl2Robots}
\begin{figure*}
\centering
\renewcommand{\figwid}{0.4\columnwidth}
{\begin{overpic}[width =\figwid]{S1_1.png}\put(10,15){$t$ = 0 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S1_2.png}\put(10,15){$t$ = 0.14 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S1_3.png}\put(10,15){$t$  = 0.29 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S1_4.png}\put(10,15){$t$  = 0.43 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S1_5.png}\put(10,15){$t$  = 0.76 s}
\end{overpic}}\\

\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:PosControl2Robots}: two robot positioning using walls with infinite friction. The algorithm only requires friction along the bottom and right walls.
Robot initial positions are shown by a crosshair, and final positions by a circled crosshair.  Dashed lines show the shortest route if robots could be controlled independently.  Solid arrows show path given by  Alg.\ \ref{alg:PosControl2Robots}.
Online demonstration and source code at \citep{Shahrokhi2015mathematicaParticle}.
%The bottom row shows an extreme case where the robots must switch position.
}
%\vspace{-2em}
}
\end{figure*}

Alg.~\ref{alg:PosControl2Robots} uses wall-friction to arbitrarily position two robots in a rectangular workspace.  This algorithm  introduces concepts that will be used for multi-robot positioning. Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.

Assume two robots are initialized at $s_1$ and $s_2$ with corresponding goal destinations $e_1$ and $e_2$. 
Denote the current positions of the robots  $r_1$ and $r_2$. 
Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $s_1.x$ and $s_1.y$ denote the $x$ and $y$ locations of $s_1$. Also the sign function is:
\begin{align}
\sign (x) = \begin{cases}  1, &  x > 0 \\
0, & x=0\\
-1, & x<0 \end{cases} 
\end{align}
The algorithm assigns a global control input at every instance.
The goal is to adjust 
 $\Delta r_x = r_2.x-r_1.x$ from $\Delta s_x = s_2.x-s_1.x$ to $\Delta e_x = e_2.x-e_1.x$ and  adjust 
 $\Delta r_y = r_2.y-r_1.y$ from $\Delta s_y = s_2.y-s_1.y$ to $\Delta e_y = e_2.y-e_1.y$ using a shared global control input. 
 This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
 %employing the assumption we have made earlier about the walls' friction.

Our algorithm solves the positioning problem in four steps: 
First, it adjusts $\Delta r_y , \Delta r_x$ as much as possible with the left wall.
Second, $\Delta r_x -\Delta e_x$ is reduced to zero with the bottom wall.
Third, if the robots were not correctly positioned relative to each other, $\Delta r_y -\Delta e_y$ is reduced to zero with the right wall.
%First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
%Second, $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant. % in Alg.~\ref{alg:YControl}. 
Lastly, the robots, now correctly positioned relative to each other, are moved to their goal locations.

In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ needs all four steps. The worst case path length is $2(\sqrt{2}+1)L$.

\begin{algorithm}
\caption{GenerateDesiredSpacing($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
\begin{algorithmic}[1]
\scriptsize
\Require knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is leftmost robot, 
 $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
 Assume $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$. If not, rotate workspace coordinates $90^{\circ}$.
\Ensure $(e_1, e_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
\State $\epsilon \gets$ robot diameter
\State $\Delta e_y = e_2.y - e_1.y$
\State $\Delta e_x = e_2.x - e_1.x$
\State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}
%Second Move
%(-\min(|-r_2.x+ \Delta e_x|, r_2.x),
\State Move $ (r_2.x, |\sign(\Delta e_y)| \, \min(r_1.y - r_2.y + \Delta e_y-\epsilon,L-r_2.y-\epsilon)+ \epsilon)$
\Statex \Comment{Adjust  $y$}
%Third Move
\State Move $(\max(\epsilon, -\sign(r_2.y - r_1.y)\Delta e_x), -\min(r_1.y , r_2.y))$ 
\Statex \Comment{Touch bottom wall}
%Fourth Move
\State Move $(\sign(r_2.y - r_1.y)\Delta e_x , (|\sign(\Delta e_y)|-1) |r_2.y - r_1.y|)$ \Comment{Adjust $x$}
\If {$\Delta e_y \neq r_2.y - r_1.y$}
%Fifth Move
%TODO This step is not correct.
\State Move $(L- \max(r_1.x, r_2.x) , \epsilon)$ \Comment{Touch right wall}
%Sixth Move
\State Move $(0, -\sign(\Delta e_x)\Delta e_y -\max(r_1.y, r_2.y)- \min(r_1.y, r_2.y)$ 
\Statex \Comment{Adjust $y$}
\EndIf
%Seventh Move
\State Move $(e_2.x - r_2.x, e_2.y - r_2.y)$ \Comment{Go to goal}
\end{algorithmic}
\end{algorithm}

\subsection{Position Control of $n$ Robots Using Wall Friction}\label{sec:PostionControlnRobots}
Alg. \ref{alg:PosControl2Robots}  can be extended to control the position of $n$ robots using wall friction under several constraints. The solution described here is an iterative procedure with $n$ loops. The $k$th loop moves the $k$th robot from a \emph{staging zone} to the desired position in a \emph{build zone}. All robots move according to the global input, but due to wall friction, at the end the $k$th loop, robots 1 through $k$ are in their desired final configuration in the build zone, and robots $k+1$ to $n$ are in the staging zone. See Fig.~\ref{fig:construction2d} for a schematic of the build and staging zones.

\begin{figure}
\begin{center}
	\includegraphics[width=1.0\columnwidth]{multirobotSliderHardware.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:construction2d}
Illustration of Alg.\ \ref{alg:PosControlNRobots}, $n$ robot position control  using wall friction.
}
\end{figure}

Assume an open workspace with four axis-aligned walls with infinite friction.
The axis-aligned build zone of dimension $(w_b, h_b)$ containing the final configuration of $n$ robots must be disjoint from the axis-aligned staging zone of dimension $(w_s, h_s)$  containing the starting configuration of $n$ robots. Without loss of generality, assume the build zone  is above the staging zone. 
Furthermore, there must be at least $\epsilon$ space above the build zone, $\epsilon$ below the staging zone, and $\epsilon + 2r$ to the left of the build and staging zone, where $r$ is the radius of a robot.  The minimum workspace is then $(\epsilon + 2r + \max(w_b,w_s), 2\epsilon + h_s,h_b)$.

The $n$ robot position control algorithm relies on a $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ control input, shown in Fig.\  \ref{fig:driftmove}.
A drift move consists of repeating a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$. The robot touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.

\begin{figure}
\begin{center}
	\includegraphics[width=.9\columnwidth]{driftmove.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:driftmove}
A  $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ to the right repeats a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$. Robot $A$ touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.}
\end{figure}

Let $(0,0)$ be the lower left corner of the workspace, $p_k$ the $x,y$ position of the $k$th robot, and $f_k$ the final $x,y$ position of the $k$th robot. Label the robots in the staging zone from left-to-right and bottom-to-top, and the $f_k$ configurations top-to-bottom and right-to-left as shown in Fig.~\ref{fig:construction2d}.

\begin{algorithm}
\caption{PositionControl$n$RobotsUsingWallFriction($k$)}\label{alg:PosControlNRobots}
\begin{algorithmic}[1]
\State move( $-\epsilon, r-p_{ky}$) % move  away from right wall and down till robot k touches bottom

\While{ $p_{kx} > r$} 
\State $\operatorname{DriftMove}(\epsilon, \min(p_{kx} - r,\epsilon), \epsilon)$ left   %drift move left until kth robot touches left wall
\EndWhile

\State $m \gets \operatorname{ceil}(\frac{f_{ky}-r}{\epsilon})$
\State $\beta \gets \frac{f_{ky}-r}{m}$
\State $\alpha \gets \beta - \frac{r - p_{ky}-\epsilon}{m}$
\For{ $m$ iterations}
\State $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ up   %move kth robot to f_{ky} and leave the rest in position.
\EndFor

\State Move ($r+\epsilon-f_{kx}, 0$)  % move the group to the left until k is in the correct relative x position
\State Move ($f_{kx}-r, 0$)

\end{algorithmic}
\end{algorithm}

Alg. \ref{alg:PosControlNRobots} proceeds as follows:  
First, the robots are moved left away from the right wall, and down so robot $k$ touches the bottom wall.
Second, a set of $\operatorname{DriftMove}$s are executed that move robot $k$ to the left wall with no net movement of the other robots.
Third, a set of $\operatorname{DriftMove}$s are executed that  move robot $k$ to its target height and return the other robots to their initial heights. 
Fourth, all robots except robot $k$ are pushed left until robot $k$ is in the correct relative $x$ position compared to robots 1 to $k-1$.
Finally, all robots are moved right until robot $k$ is in the desired target position. Running time is $O(n(w+h))$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Covariance Control}
\label{sec:covControl}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htb]
\begin{center}
\includegraphics[width=\linewidth]{SquareFill.pdf} 
\vspace{-1em}
\caption{Pushing the swarm against a square boundary wall allows limited control of the shape of the swarm, as a function of swarm area $A$ and the commanded movement direction $\beta$. Left plot shows locus of possible mean positions for five values of $A$.  The locus morphs from a square to a circle as $A$ increases.  The covariance ellipse for each $A$ is shown with a dashed line. Center shows two corresponding arrangements of kilobots.  At right is $\bar{x}(A), \sigma_{xy}(A), \sigma_x^2(A),$ and $\rho(A)$ for a range of $\beta$ values. See online interactive demonstration at \citep{Zhao2016mathematicaSquare}.}
\label{fig:SquareFill}
\end{center}
\end{figure*} 
\begin{figure*}[!htb]
\begin{center}
\vspace{-1em}
\includegraphics[width=\linewidth]{CircleFill.pdf} 
\vspace{-1em}
\caption{Pushing the swarm against a circular boundary wall allows limited control of the shape of the swarm, as a function of the fill level $h$ and the commanded movement direction $\beta$. Left plot shows locus of possible mean positions for four values of $h$. The locus of possible mean positions are concentric circles. See online interactive demonstration at \citep{Zhao2016mathematica}.}  
\label{fig:CircleFill}
\end{center}
\vspace{-1em}
\end{figure*}

@@

The next section shows how a system in a rectangularly bounded workspace with friction model \eqref{eq:frictionmodel} can arbitrarily position two robots. 
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\columnwidth]{friction.pdf} 
\vspace{-1em}
\caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (c) velocity of a fluid reduces to zero at the boundary.
\label{fig:friction}
}\vspace{-1em}
\end{center}
\end{figure}

### 2018-06-11 Shiva Dissertation/chap2/Introduction.tex
Shiva PhD dissertation

%Micro- and nanorobotics have the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.  The same properties that promise breakthrough solutions---small size and large populations---present unique challenges to generating controlled motion.  
Large populations of micro- and nanorobots are being produced in laboratories around the world, with diverse potential applications in drug delivery and construction, see \cite{Peyer2013,Shirai2005,Chiang2011}. These activities require robots that behave intelligently.
Limited computation and communication at small scales makes autonomous operation or direct control over individual robots difficult. 
Instead, this paper treats the robots as particles that are steered by a global control signal broadcast to the entire population.    
This paper examines object manipulation by a swarm of particles, as illustrated in Fig.~{\ref{fig:bigPictureMeanAndVarianceForSwarm}. 
The transportation methodology is similar to that in~\cite{sugawara2014object}, but rather than using onboard computation or sensing, the particles all move in the same direction.

@@

Many promising applications for particle swarms require direct human control, but user interfaces to thousands---or millions---of particles is a daunting human-swarm interaction (HSI) challenge. Our early work with over a hundred hardware robots and thousands of simulated particles demonstrated that direct human control of large swarms is possible, \cite{Becker2013b}. 
Unfortunately, the logistical challenges of repeated experiments with over a hundred robots prevented large-scale tests. 
There is currently no comprehensive understanding of user interfaces for controlling multi-robot systems with massive populations.  
One contribution of this paper is a tool for investigating HSI methods through statistically significant numbers of experiments.

\begin{figure}
\centering
\begin{overpic}[width=\columnwidth]{MainExpFig.pdf}\end{overpic}
\caption{\label{fig:bigPictureMeanAndVarianceForSwarm} 
A swarm of particles, all actuated by a uniform control input where each particle gets the same control input, can be effectively manipulated by a control law that uses only the mean and variance of the robot distribution.  
Here a swarm of particles (kilobot robots) pushes a green hexagon toward the goal (see video attachment).
}
\end{figure}

%Inspired by our online experiments, because it is not always possible to gather pose information on each robot for feedback control and
%It is not always practical to gather pose information on individual robots for feedback control; the robots might be difficult or impossible to sense individually due to their size and location. However, it is often possible to sense global properties of the group, such as mean position and density.
Often particles are difficult or impossible to sense individually due to their size and location. 
For example, microrobots are smaller than the minimum resolution of a clinical MRI-scanner, see \cite{martel2014computer}, however it is often possible to sense global properties of the group such as mean position and variance. 
To make progress in automatic control with global inputs, this paper presents swarm manipulation controllers inspired by our online experiments that require only mean and variance measurements of the particle's positions. 
 To perform the object manipulation task illustrated in Fig.~\ref{fig:bigPictureMeanAndVarianceForSwarm}, we use these controllers   as primitives, policy iteration for path planning, handle outliers by partitioning the workspace, and minimize pushing the object backwards with potential field navigation.

 %function and implementation function

 Our paper is organized as follows.  After a discussion of related work in \S \ref{sec:RelatedWork},  we describe our experimental methods for an online human-user experiment and their results in \S \ref{sec:expMethods}. Next we prove that the mean and variance of a particle swarm are controllable in \S \ref{sec:theory}, and present automatic controllers in \S \ref{sec:simulation}. We use these controllers as primitives and present a framework for manipulating an object through a maze in \S \ref{sec:exp}. 
 We implement these controllers in our hardware robots and use them to complete an object manipulation task with 100 kilobots in \S \ref{sec:realExperiment}, and conclude in \S \ref{sec:conclusion}.

>>>>

%Micro- and nanorobotics have the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.  The same properties that promise breakthrough solutions---small size and large populations---present unique challenges to generating controlled motion.  
Large populations of micro- and nanorobots are being produced in laboratories around the world, with diverse potential applications in drug delivery and construction, see \cite{Peyer2013,Shirai2005,Chiang2011}. %These activities require robots that behave intelligently.
Limited computation and communication at small scales makes autonomous operation or direct control over individual robots difficult. 
Instead, this chapter treats the robots as particles that are steered by a global control signal broadcast to the entire population.    
This chapter examines object manipulation by a swarm of particles, as illustrated in Fig.~{\ref{fig:bigPictureMeanAndVarianceForSwarm}. 
The transportation methodology is similar to that in~\cite{sugawara2014object}, but rather than using onboard computation or sensing, the particles all move in the same direction.

@@

Many promising applications for particle swarms require direct human control, but user interfaces to thousands---or millions---of particles is a daunting human-swarm interaction (HSI) challenge. Human control of large swarms is possible over a hundred hardware robots and thousands of simulated particles \cite{Becker2013b}. 
%Unfortunately, the logistical challenges of repeated experiments with over one hundred robots prevented large-scale tests. 
%There is currently no comprehensive understanding of user interfaces for controlling multi-robot systems with massive populations.  
%One contribution of this paper is a tool for investigating HSI methods through statistically significant numbers of experiments.

\begin{figure}
\centering
\begin{overpic}[width=\columnwidth]{MainExpFig.pdf}\end{overpic}
\caption{\label{fig:bigPictureMeanAndVarianceForSwarm} 
A swarm of particles, all actuated by a uniform control input %where each particle gets the same control input, 
can be effectively manipulated by a control law that uses only the mean and variance of the robot distribution.  
Here a swarm of kilobots pushes a green hexagon toward the goal.
}
\end{figure}

%Inspired by our online experiments, because it is not always possible to gather pose information on each robot for feedback control and
%It is not always practical to gather pose information on individual robots for feedback control; the robots might be difficult or impossible to sense individually due to their size and location. However, it is often possible to sense global properties of the group, such as mean position and density.
%Often particles are difficult or impossible to sense individually due to their size and location. 
%For example, microrobots are smaller than the minimum resolution of a clinical MRI-scanner, see \cite{martel2014computer}, however it is often possible to sense global properties of the group such as mean position and variance. 
 This chapter presents swarm manipulation controllers inspired by our online experiments that require only mean and variance measurements of the positions of the particles. 
 To perform the object manipulation task illustrated in Fig.~\ref{fig:bigPictureMeanAndVarianceForSwarm}, we use these controllers  as primitives, policy iteration for path planning, handle outliers by partitioning the workspace, and minimize pushing the object backwards with potential field navigation.

 %function and implementation function
This chapter is an extended version of \cite{shahrokhi2018tro} and is organized as follows.  After a discussion of related work in Sec. \ref{sec:RelatedWork},  we describe our experimental methods for an online human-user experiment and their results in Sec. \ref{sec:expMethods}. Next we prove that the mean and variance of a particle swarm are controllable in Sec. \ref{sec:theory}, and present automatic controllers in Sec. \ref{sec:simulation}. We use these controllers as primitives and present a framework for manipulating an object through a maze in Sec. \ref{sec:exp}. 
 We implement these controllers in our hardware robots and use them to complete an object manipulation task with 100 kilobots in Sec. \ref{sec:realExperiment}, and conclude in Sec. \ref{sec:conclusion}.

### 2018-06-11 Shiva Dissertation/chap2/conclusion.tex
Shiva PhD dissertation

    %Micro- and nanorobotics have the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.  
  The small size of micro and nano particles makes individual control and autonomy challenging, so currently these particles are steered by global control inputs such as magnetic fields or chemical gradients. To investigate this control challenge, this paper introduced \href{http://www.swarmcontrol.net}{SwarmControl.net}, an open-source tool for large-scale user experiments where human users steer swarms of robots to accomplish tasks.  Analysis of the gameplay results revealed benefits of measuring and controlling statistics of the swarm rather than full state feedback, robustness to IID noise, and small effects of varying population size of large swarms.

@@

Our future goal is to perform assembly using particle swarms to manipulate and attach components. This task requires applying force and torque to components and manipulating them through obstacles and each other. This work provides foundational algorithms and techniques for steering swarms, object manipulation, and addressing obstacle fields, but there are many opportunities to extend the work.

Topics of interest include control with nonuniform flow such as fluid in an artery, gradient control fields like that of an MRI, competitive playing, multi-modal control, flexible workspaces, optimal-control, and targeted drug delivery in a vascular network.

>>>>

    %Micro- and nanorobotics have the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.  
  The small size of micro and nano particles makes individual control and autonomy challenging, so currently these particles are steered by global control inputs such as magnetic fields or chemical gradients. To investigate this control challenge, this chapter introduced \href{http://www.swarmcontrol.net}{SwarmControl.net}, an open-source tool for large-scale user experiments where human users steer swarms of robots to accomplish tasks.  Analysis of the gameplay results revealed benefits of measuring and controlling statistics of the swarm rather than full state feedback, robustness to IID noise, and small effects of varying population size of large swarms.

@@

Our goal in the next chapter is to apply force and torque to components and manipulating them through obstacles and each other. This work provides foundational algorithms and techniques for steering swarms, object manipulation, and addressing obstacle fields.%, but there are many opportunities to extend the work.

%Topics of interest include control with nonuniform flow such as fluid in an artery, gradient control fields like that of an MRI, competitive playing, multi-modal control, flexible workspaces, optimal-control, and targeted drug delivery in a vascular network.

### 2018-06-11 Shiva Dissertation/chap2/experiment.tex
Shiva PhD dissertation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Particle Swarm Object manipulation}\label{sec:exp}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section analyzes an \emph{object manipulation} task attempted by our hybrid, hysteresis-based controllers. The swarm must deliver the object to the goal region. 
%We assume that particles pushing the object move slower than unencumbered particles, and that pushing with more particles increases the object speed
We assume Coulomb and viscous friction parameters such that the object can be moved by particle motion.  Increasing the number of pushing particles increases the object speed.
%Although one particle might be able to move the object, it requires a long time for it to complete the task and probability of getting stuck when there is not enough particles is high.
 To solve this object manipulation task we divide the task into three components: 1) designing a policy for the object, 2) pushing the object with a compliant swarm, and 3) managing outliers.

The table below summarizes the simulation results for each 10 successful trials:
\begin{center}
\begin{tabular}{ c c }
\hline
Method & Result, mean$\pm$std (s)\\ [0.5ex] 
\hline \hline
Value Iteration (VI) & 367 $\pm$ 253 \\ 
\hline
VI + Potential Fields (PF) & 271 $\pm$ 267 \\  
\hline
VI + Outlier Rejection (OR) & 245 $\pm$ 135\\
\hline
BFS + PF + OR & 183 $\pm$ 179 \\
\hline
VI + PF + OR & 90 $\pm$ 35 \\[0.5ex]
\hline
\end{tabular}
\end{center}
\subsection{Learning a policy for the object}\label{subsec:objectpolicy}

@@

\begin{figure}
\centering
\renewcommand{\figwid}{3cm}
\begin{overpic}[height=\figwid]{GradientView.png}
\end{overpic}
\begin{overpic}[height=\figwid]{PolicyIter.png}
\end{overpic}
\begin{overpic}[height=\figwid]{MDPmap.pdf}
\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:BFSGradient} BFS finds the shortest path for the moveable object to compute gradient vectors (left). Modeling as an MDP enables encoding penalties for being near obstacles. (Middle) The control policy from value iteration. (Right) The vision algorithm detects obstacles in the hardware setup. This map is used to produce the value function and control policy shown. 
%\vspace{-2em}
}
\end{figure}

@@

\begin{figure}\begin{center}\begin{overpic}[width=0.4\columnwidth]{MDPEmaze.pdf}\end{overpic}~\begin{overpic}[width=0.4\columnwidth]{EmazeFinal.pdf}\end{overpic}\\
\vspace{.25em}
\begin{overpic}[width=0.4\columnwidth]{Spiral.pdf}\end{overpic}~\begin{overpic}[width=0.4\columnwidth]{SpiralSuccess.pdf}\end{overpic}
\end{center}
\caption{\label{fig:Emaze} We tested three workspaces. The control policies from value iteration for an {\sffamily E}-shaped and a spiral workspace are shown in the left column.
}
\end{figure}
\begin{figure}
\centering
\begin{overpic}[width=.5\columnwidth]{DifferentShapes.pdf}\end{overpic}
\vspace{-1em}
\caption{\label{fig:DifferentMazes} Completion times for three workspaces.}
\end{figure}

@@

\begin{figure*}
\centering
\renewcommand{\figwid}{0.5\columnwidth}
\begin{overpic}[width =\figwid]{SimVeryNum.pdf}\put(1,55){a)}
\end{overpic}
\begin{overpic}[width =\figwid]{SimVeryNoise.pdf}\put(1,55){b)}
\end{overpic}
\begin{overpic}[width =\figwid]{SimVeryWeight.pdf}\put(1,55){c)}
\end{overpic}
\begin{overpic}[width =\figwid]{SimVeryShape.pdf}\put(1,55){d)}
\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:AutoVeryParam}Parameter sweep simulation studies for a) number of particles, b) different noise values, c) object weight, and d) object shape.  Each bar is labelled with the number of trials. Completion time is in seconds.
%\vspace{-2em}
}
\end{figure*}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Particle swarm object manipulation}\label{sec:exp}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\vspace{10em}
\begin{table}[h!]
  \begin{center}
  %\vspace{-1em}
    \caption{\label{tbl:resultsSum}Summary, Object Manipulation Results}
\begin{tabular}{ c c }
\hline
Method & Result (10 trials), mean$\pm$std (s)\\ [0.5ex] 
\hline \hline
Value Iteration (VI) & 367 $\pm$ 253 \\ 
\hline
VI + Potential Fields (PF) & 271 $\pm$ 267 \\  
\hline
VI + Outlier Rejection (OR) & 245 $\pm$ 135\\
\hline
BFS + PF + OR & 183 $\pm$ 179 \\
\hline
VI + PF + OR & 90 $\pm$ 35 \\[0.5ex]
\hline
\end{tabular}
\end{center}
%\vspace{-1em}
\end{table}

This section analyzes an \emph{object manipulation} task attempted by our hybrid, hysteresis-based controllers. The swarm must deliver the object to the goal region. 
%We assume that particles pushing the object move slower than unencumbered particles, and that pushing with more particles increases the object speed
We assume Coulomb and viscous friction parameters such that the object can be moved by particle motion.  Increasing the number of pushing particles increases the object speed.
%Although one particle might be able to move the object, it requires a long time for it to complete the task and probability of getting stuck when there is not enough particles is high.
 To solve this object manipulation task we divide the task into three components: 1) designing a policy for the object, 2) pushing the object with a compliant swarm, and 3) managing outliers. Table \ref{tbl:resultsSum} summarizes results for successful simulation trials.

\subsection{Learning a policy for the object}\label{subsec:objectpolicy}

@@

\begin{figure}
\centering
%\renewcommand{\figwid}{5cm}
\begin{overpic}[width=0.33\columnwidth]{GradientView.png}
\end{overpic}
\begin{overpic}[width=0.33\columnwidth]{PolicyIter.png}
\end{overpic}
\begin{overpic}[width=0.255\columnwidth]{MDPmap.pdf}
\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:BFSGradient} BFS finds the shortest path for the moveable object to compute gradient vectors (left). Modeling as an MDP enables encoding penalties for being near obstacles. (Middle) The control policy from value iteration. (Right) The vision algorithm detects obstacles in the hardware setup. %This map is used to produce the value function and control policy shown. 
%\vspace{-2em}
}
\end{figure}

@@

\begin{figure}\begin{center}\begin{overpic}[width=0.4\columnwidth]{MDPEmaze.pdf}\end{overpic}~\begin{overpic}[width=0.4\columnwidth]{EmazeFinal.pdf}\end{overpic}\\
\vspace{.25em}
\begin{overpic}[width=0.4\columnwidth]{Spiral.pdf}\end{overpic}~\begin{overpic}[width=0.4\columnwidth]{SpiralSuccess.pdf}\end{overpic}
\end{center}
\caption{\label{fig:Emaze} We tested three workspaces. The control policies from value iteration for an {\sffamily E}-shaped and a spiral workspace are shown in the left column.
}
\end{figure}
\begin{figure}
\centering
\begin{overpic}[width=\columnwidth]{DifferentShapes.pdf}\end{overpic}
\vspace{-1em}
\caption{\label{fig:DifferentMazes} Completion times for three workspaces.}
\end{figure}

@@

\begin{figure*}
\centering
\renewcommand{\figwid}{0.49\columnwidth}
\begin{overpic}[width =\figwid]{SimVeryNum.pdf}\put(1,55){a)}
\end{overpic}
\begin{overpic}[width =\figwid]{SimVeryNoise.pdf}\put(1,55){b)}
\end{overpic}\\
\vspace{2em}
\begin{overpic}[width =\figwid]{SimVeryWeight.pdf}\put(1,55){c)}
\end{overpic}
\begin{overpic}[width =\figwid]{SimVeryShape.pdf}\put(1,55){d)}
\end{overpic}
%\vspace{-0.5em}
\caption{\label{fig:AutoVeryParam}Parameter sweep simulation studies for a) number of particles, b) different noise values, c) object weight, and d) object shape.  Each bar is labelled with the number of trials. Completion time is in seconds.
%\vspace{-2em}
}
\end{figure*}
